---
title: "Retreat_Discharge"
output: html_document
date: "2024-09-23"
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to take velocity transect measurements collected from the Swoffer Wading rod and convert to discharge.

Before running this script workflow:

Step 1) update "SSF_Swoffer_Metadata" meta data sheet with values from the field 
  *Make sure units are properly labelled. 

Script Workflow:

Step 1) Load in "SSF_Swoffer_Metadata" meta data sheet from googledrive located in 

Step 2) run for-loop chunk that calculates discharge for each site 

Step 3) Save output file as: in "enter_directory filepath"

### Load in Libraries ###
```{r}
rm(list = ls())

library(tidyverse)
library(gsheet)
library(zoo)
library(dplyr)
```

# test #
```{r}
# Q <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1sr8z7Jn-_n3M3T3se8LL1ZOSHFZhrCyDL1aEax3aT38/edit#gid=0')
Q <- read_csv("~/Documents/RC_3/Fire_Watch/2024/Swoffer/oak_creek_swoffer.csv") %>% 
  mutate(Date = mdy(Date))
###### TEST WITH THE DIFFERENT UNITS - 
# **if the unit is "m" than multiply by 100
# **if the unit is "cm" than print
# **if the unit is "ft" than multiply by 30.48 to all get it to cm's 

Q <- Q %>% 
  mutate(Water_Depth_cm = case_when(Water_Depth_unit == 'm' ~ Water_Depth * 100,
                                    Water_Depth_unit == 'ft' ~ Water_Depth * 30.48,
                                    Water_Depth_unit == 'cm' ~ print(Water_Depth)),
         
         Distance_cm = case_when(Distance_unit == 'm' ~ Distance * 100,
                                    Distance_unit == 'ft' ~ Distance * 30.48,
                                    Distance_unit == 'cm' ~ print(Distance)))


# BRIE for loop test # 
formatted <- Q %>% 
  mutate(site_DateTime = str_c(Site_ID, Date, Start_Time, sep = "_"),
         DateTimePST = ymd_hms(paste(Date, hms(Start_Time))))

# creating an empty tibble that will be populated with the for-loop data 
combine <- tibble(Site_ID = as.character(), 
                  DateTimePST = as.character(),
                  FinalQ = as.numeric())

# list of what the for loop is looping through (site_DateTime from the "formatted" df)
sites <- unique(formatted$site_DateTime)

# for loop #
# variable - something that doesnt exist yet and what you will be creating

for (site in sites) {
  
  filter <- formatted %>% 
    filter(site_DateTime == site) %>% 
    mutate(VeloMean = rowMeans(dplyr::select(., starts_with("Velocity")), na.rm = TRUE))
  
  if(nrow(filter) > 2) {
    
    summary <- filter %>% 
      dplyr::group_by(site_DateTime, Point) %>% 
      summarise(Site_ID = unique(Site_ID),
                DateTimePST = unique(DateTimePST),
                Water_Depth_cm = unique(Water_Depth_cm),
                Distance_cm = unique(Distance_cm),
                VeloMeanMean = mean(VeloMean, na.rm = TRUE)) %>% 
      mutate(VeloMeanPoints = rollapply(VeloMeanMean, 2, mean, align = 'right', fill = NA),
           DistDiff = rollapply(Distance_cm, 2, diff, align = 'right', fill = NA),
           DepthMean = rollapply(Water_Depth_cm, 2, mean, align = 'right', fill = NA),
           TransectQ = ((DistDiff * DepthMean)/10000000) * (VeloMeanPoints * 1000),
           FinalQ = sum(TransectQ, na.rm = TRUE)) %>% 
      ungroup() %>% 
      dplyr::select(Site_ID, DateTimePST, FinalQ) %>% 
      distinct() %>% 
      mutate(DateTimePST = paste(" ", as.character(DateTimePST), sep = ""))
    
    combine <- combine %>% 
    add_row(summary)
  
  }
  
}

write_csv(formatted, "~/Documents/RC_3/Fire_Watch/2024/Swoffer/Oak_Creek_Swoffer.csv")




```



### PLOT ###
```{r}
ggplot(combine) +
  geom_point(aes(x = as.Date(DateTimePST), y = FinalQ, color = Site_ID), size = 3) +
  ylab("Discharge (m^3/s)") +
  xlab("Date") +
  facet_wrap(~Site_ID, scales = "free") +
  theme_bw() 
  

ggsave("Q.figure.pdf",
       path = ("~/Volumes/sbr_sfa/RC3/00_Schneider_Springs_Fire_2023/SSF_Data_Processing/codes"),
       width = 10, height = 8, units = "in")


ggplot(combine) +
  geom_point(aes(x = as.Date(DateTimePST), y = FinalQ, color = Site_ID), size = 3) +
  ylab("Discharge (m^3/s)") +
  xlab("Date") +
  theme_bw() 

# Plot cross section depth
ggplot(Q) +
  geom_point(aes(x = Point, y = Water_Depth_cm)) +
  geom_hline(yintercept = 0, color = "blue") +
  scale_y_reverse() +
  theme_bw()

# Liters per second
combine$FinalQ*1000 # 31.65869

# cfs 
combine$FinalQ*35.314666212661 # 1.118016


```






# old script #
```{r}
S02 <- filter(Q, Site_ID == "S02" & Start_Time == "7:45")

# Concatenate date & time
S02$DateTimePST <- as.POSIXct(paste(S02$Date, S02$Start_Time), format = "%Y-%m-%d %H:%M")

S02 <- S02 %>% 
  mutate(VeloMean = rowMeans(select(., starts_with("Velocity")), na.rm = TRUE),
         VeloMeanPoints = rollapply(VeloMean, 2, mean, align = 'right', fill = NA),
         DistDiff = rollapply(Distance_cm, 2, diff, align = 'right', fill = NA),
        
         DepthMean = rollapply(Water_Depth_cm, 2, mean, align = 'right', fill = NA),
         TransectQ = ((DistDiff * DepthMean)/10000) * (VeloMeanPoints * 1000),
         FinalQ = sum(TransectQ, na.rm = TRUE))


```


